### 2022-06-27 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@kazewong}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit. 
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis. 
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna. 
    Donec vehicula augue eu neque, morbi tristique senectus et netus et. 
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum. 
    Phasellus eu tellus sit amet tortor gravida placerat. 
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. 
    Praesent eget sem vel leo ultrices bibendum. 
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac. 
    Curabitur auctor semper nulla donec varius orci eget risus. 
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam. 
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. 
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis. 
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna. 
Donec vehicula augue eu neque, morbi tristique senectus et netus et. 
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum. 
Phasellus eu tellus sit amet tortor gravida placerat. 
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. 
Praesent eget sem vel leo ultrices bibendum. 
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac. 
Curabitur auctor semper nulla donec varius orci eget risus. 
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam. 
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi. 
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae. 
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna. 
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante. 
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient. 
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis. 
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2022-06-27 src/tex/ms.tex
add files. Somehow bib doesn't compile

@@
-% Define document class
-\documentclass[twocolumn]{aastex631}
+%%%%%%%% ICML 2022 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
+
+\documentclass[nohyperref]{article}
 \usepackage{showyourwork}
 
-% Begin!
+% Recommended, but optional, packages for figures and better typesetting:
+\usepackage{microtype}
+\usepackage{graphicx}
+\usepackage{subfigure}
+\usepackage{booktabs} % for professional tables
+
+% hyperref makes hyperlinks in the resulting PDF.
+% If your build breaks (sometimes temporarily if a hyperlink spans a page)
+% please comment out the following usepackage line and replace
+% \usepackage{icml2022} with \usepackage[nohyperref]{icml2022} above.
+\usepackage{hyperref}
+\usepackage{xspace} % Space only when needed.
+\usepackage{xcolor}
+\definecolor{rb4}{HTML}{27408B}
+\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
+\newcommand{\mc}[1]{{\color{purple}[MC: #1 ]}}
+\newcommand\pysr{\textit{PySR}\xspace}
+
+% Attempt to make hyperref and algorithmic work together better:
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
+
+% Use the following line for the initial blind version submitted for review:
+\usepackage{icml2022}
+
+% If accepted, instead use the following line for the camera-ready submission:
+% \usepackage[accepted]{icml2022}
+
+% For theorems and such
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{mathtools}
+\usepackage{amsthm}
+
+% if you use cleveref..
+\usepackage[capitalize,noabbrev]{cleveref}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% THEOREMS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{assumption}[theorem]{Assumption}
+\theoremstyle{remark}
+\newtheorem{remark}[theorem]{Remark}
+
+% Todonotes is useful during development; simply uncomment the next line
+%    and comment out the line below the next line to turn off comments
+%\usepackage[disable,textsize=tiny]{todonotes}
+\usepackage[textsize=tiny]{todonotes}
+
+% The \icmltitle you define below is probably too long as a header.
+% Therefore, a short form for the running title is supplied here:
+%\icmltitlerunning{Automated discovery of interpretable gravitational-wave population models}
+
 \begin{document}
 
-% Title
-\title{An open source scientific article}
+\twocolumn[
+%\icmltitle{Automatic, Flexible and Interpretable:\\gravitational-wave population analysis with symbolic regression}
+%\icmltitle{Automated discovery of symbolic gravitational-wave population models}
+\icmltitle{Automated discovery of interpretable gravitational-wave population models}
 
-% Author list
-\author{@kazewong}
+\icmlsetsymbol{equal}{*}
 
-% Abstract with filler text
+\begin{icmlauthorlist}
+\icmlauthor{Kaze W.K Wong}{equal,yyy}
+\icmlauthor{Miles Cranmer}{equal,yyy,comp}
+\end{icmlauthorlist}
+
+\icmlaffiliation{flatiron}{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
+
+\icmlcorrespondingauthor{Kaze Wong}{kwong@flatironinstitute.org}
+
+% You may provide any keywords that you
+% find helpful for describing your paper; these are used to populate
+% the "keywords" metadata in the PDF but will not be shown in the document
+\icmlkeywords{Machine Learning, ICML}
+
+\vskip 0.3in
+]
+
+% this must go after the closing bracket ] following \twocolumn[ ...
+
+% This command actually creates the footnote in the first column
+% listing the affiliations and the copyright notice.
+% The command takes one argument, which is text to display at the start of the footnote.
+% The \icmlEqualContribution command is standard text for equal contribution.
+% Remove it (just {}) if you do not need this facility.
+
+%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
+\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
+
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit. 
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis. 
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna. 
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et. 
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum. 
-    Phasellus eu tellus sit amet tortor gravida placerat. 
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. 
-    Praesent eget sem vel leo ultrices bibendum. 
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac. 
-    Curabitur auctor semper nulla donec varius orci eget risus. 
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam. 
-    Duis eget orci sit amet orci dignissim rutrum.
+We present an approach to automatically discover analytic population models for gravitational-wave (GW) events from data.
+As more gravitational-wave (GW) events are detected, flexible models such as Gaussian Mixture Models have become more important in fitting the distribution of GW properties due to their expressivity.
+However, flexible models come with a cost: a large number of parameters that lack physical motivation, making interpreting the implication of these models very difficult.
+In this work, we demonstrate the use of symbolic regression to distill such flexible models into interpretable analytic expressions.
+We recover common GW population models such as a power-law-plus-Gaussian, and find a new empirical population model which combines accuracy and simplicity.
+This demonstrates a promising pathway to automatically discover interpretable descriptions of features in the ever-growing GW catalog. 
 \end{abstract}
 
-% Main body with filler text
 \section{Introduction}
-\label{sec:intro}
+% \begin{figure*}[hbt!]
+% % \includegraphics[width=\textwidth]{figures/GWTC3_m1_combined.pdf}
+% \includegraphics[width=\textwidth]{figures/Page2.pdf}
+% \caption{An illustration of our proposed pipeline, which first
+% (\textit{far left})
+% requires observations of individual source properties (here, events shown from GWTC3);
+% then
+% (\textit{middle left}) 
+% infers the deconvolved population posterior density;
+% followed by (\textit{middle right})
+% using symbolic regression to search for an
+% an analytic model which approximates the density, using
+% operators common to existing population models; and finally
+% (\textit{far right}) distilling physical insights
+% from the recovered population model.
+% }
+% %\caption{\textit{left}: posterior density of the primary mass from all events reported in GWTC3. Each line is the normalized posterior density for one event.
+% %\textit{middle}: posterior predictive distribution of the Gaussian mixture model reported in \cite{LIGOScientific:2021psn}.
+% %\textit{right}: equations tabulated in table \ref{tab:equations}}
+% \label{fig:Gaussian_m1_fit}
+% \end{figure*}
+The rate of gravitational wave (GW) detections increases exponentially as the sensitivity of GW detectors improves \cite{LIGOScientific:2021djp} since the discovery in 2015 \cite{Abbott:2016blz}.
+To date, there are 79 officially announced events, the majority of them being binary black holes (BBH).
+With the increasing number of observed GW events, more and more features in the distribution of GW events await to be discovered.
+Currently, one of the most common ways to understand the observed GW population is through phenomenological modeling.
+For example, to understand the distribution of primary mass in observed BBHs, one would write down a hand-crafted model of the primary mass distribution, such as a power-law distribution \cite{LIGOScientific:2021psn}.
+By comparing the model to the data, one can measure the parameters which characterize the model, such as the spectral index of a power-law distribution.
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. 
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis. 
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna. 
-Donec vehicula augue eu neque, morbi tristique senectus et netus et. 
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum. 
-Phasellus eu tellus sit amet tortor gravida placerat. 
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. 
-Praesent eget sem vel leo ultrices bibendum. 
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac. 
-Curabitur auctor semper nulla donec varius orci eget risus. 
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam. 
-Duis eget orci sit amet orci dignissim rutrum.
+The central issue with the use of this paradigm is the lack of scalability to a more complex dataset, since manual input is required to write down a sensible phenomenological model.
+After the announcement of 11 events in the first Gravitational-Wave Transient Catalog (GWTC1) in 2019 \cite{LIGOScientific:2018mvr}, a power law was sufficient to explain the primary mass distribution.
+In GWTC2 (which is announced in 2020) \cite{LIGOScientific:2020kqk}, the community modified the simple power-law to either a broken power law with two spectral indices or with the superimposing of a Gaussian distribution on top of the power-law.
+Finally, the state-of-the-art model used in analyzing GWTC3 \cite{LIGOScientific:2021psn} further adds a spline on top of the power-law and Gaussian model to fit the residual \cite{Edelman:2021zkw}.
+Since the complexity of the catalog grows with the size of the catalog,
+the community is struggling to write down an interpretable model that is flexible enough to explain the increasing rich set of features in the catalog.
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi. 
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae. 
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna. 
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante. 
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient. 
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis. 
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+Alternatively, there have been studies using much more flexible models such as a Gaussian mixture model to analyze the catalog \cite{Tiwari:2020vym}.
+While such a flexible model is powerful in fitting complex datasets,
+it is difficult to interpret the physical meaning of the fitted parameters.
+There are two reasons why flexible models are hard to interpret:
+first, flexible models often have many parameters.
+In GWTC3, the specific Gaussian mixture model has over 100 parameters.
+With that number of parameters, it is not trivial to assess the relative importance of the parameters.
+Second, flexible models often employ bases that are not physically meaningful.
+For example, consider that one wishes to assess which parameter in the fitted Gaussian mixture model is the most important, one may apply a principal component analysis (PCA) to the set of parameters and find the most important component.
+However, the PCA basis for this dataset is not a basis that is directly related to any astrophysical process,
+so having the relative importance in that form does not provide insight into astrophysics.
 
-\bibliography{bib}
+Symbolic regression (SR) is a machine learning technique which searches to space of analytic expressions to fit a dataset---this algorithm complements the weakness of a flexible model in understanding the distribution of GW.
+Since SR does not take advantage of using gradient information when searching for expressions,
+and the space of equation is a combinatorial space,
+SR is slow in comparison to traditional machine learning techniques such as Gaussian Mixture Models (GMMs).
+Therefore, rather than fitting SR directly to the complex and expensive likelihoods found in GW event analysis, we first fit the GMM as a sort of approximated likelihood.
+Then, we apply SR to the output of this GMM, as the loss function is then simply mean-squared error compared to the GMM, rather equal to an evaluation of the raw likelihood.
+%While SR itself is not efficient enough to be used to analyze the data directly, using SR to interpret a flexible model that has already been fitted to the data can be very effective.
+
+In this work, we apply SR to the mixture Gaussian model used to fit the primary mass distribution of GWTC3.
+In particular, we apply SR to fit the posterior predictive distribution given by the Gaussian mixture model.
+This paper is structured as follows: we describe the GW data used in this work in \ref{sec:GWdata}.
+We briefly review SR in section \ref{sec:method}.
+We present our main result in section \ref{sec:result}.
+We discuss the potential usage and extension of this work in section \ref{sec:discussion}.
+
+\section{Method}
+\label{sec:method}
+
+\subsection{Symbolic regression}
+
+Symbolic regression is a machine learning (ML) algorithm for regression which searches the space of analytic expressions that fit some dataset.
+Whereas most ML algorithms are optimized by fitting a fixed set number of real parameters, symbolic regression is typically done using a ``genetic algorithm'' which finds a tree or sequence of symbols that optimizes an objective.
+Since models for physical phenomena are often specified in a language of concise mathematical expressions, symbolic regression can produce very interpretable models compared to traditional ML, as the ``trained model'' generated by symbolic regression is in the same language as existing physical models.
+Therefore, symbolic regression can produce insights for science in a way that traditional ML cannot.
+
+% Give history
+%Symbolic regression has a disconnected history,
+% with some early work going back to \mc{langley citations}.
+%The first application of genetic algorithms
+%to symbolic regression was done by \mc{koza}, before this class of algorithms was popularized in the sciences by \mc{lipson}.
+% \mc{Cite examples in astronomy.}
+% \mc{Describe how PySR works.}
+
+\begin{table*}[hbt!]
+    \begin{center}
+    \begin{tabular}{ l l l l }
+    \hline
+    \hline
+    Complexity &  Score & Loss & Equation\\
+    \hline
+    \hline
+    16 & 0.126   & 0.212  & $3.72 \rm{Cond}{\left(M - 9.27,0.9^{M} \right)} + 3.72 \rm{Gauss}{\left(0.52 M - 4.85 \right)}$\\
+    25 & 0.501   & 0.0352 & $3.89 \left(\rm{Gauss}{\left(0.19 M - 6.54 \right)} + 0.45\right) \rm{Cond}{\left(M - 9.12,0.91^{M} \right)}$\\
+    & & & $ + 3.89 \rm{Gauss}{\left(0.5 M - 4.8 \right)}$\\
+    48 & 0.00771 & 0.0108 & $\left(\rm{Cond}{\left(M - 9.42,0.62 \cdot 0.9^{M} \right)} + 1.44 \rm{Gauss}{\left(0.51 M - 4.88 \right)}\right)$\\
+    & & & $\left(5.11 \rm{Gauss}{\left(0.06 M - 4.67 \right)} + 7.82 \rm{Gauss}{\left(0.17 M - 5.86 \right)} + 3.26\right)$\\
+    \hline
+    \hline
+    \end{tabular}
+    \caption{Equations return obtained through symbolic regression with \pysr.
+    In the search we perform, there are 30 equations with different complexities.
+    We select three most representative equations using a combination of score and loss threshold.}
+    \label{tab:equations}
+    \end{center}
+\end{table*}
+
+\paragraph{Algorithm.}
+There are several different strategies for performing symbolic regression, all of them offering different advantages.
+Some methods, such as \textit{SINDy} \cite{bruntonDiscoveringGoverningEquations2016}, are dictionary-based, and search for a sparse linear combination of hand-crafted terms.
+Other methods, such as \textit{eureqa} \cite{schmidtDistillingFreeFormNatural2009}, are genetic algorithm-based, and, though slower, can find any expression that can be represented as a tree of predefined operators, variables, and constants.
+
+We use the algorithm \pysr\footnote{\url{https://github.com/MilesCranmer/PySR}} \cite{cranmerPySRFastParallelized2020} for performing symbolic regression.
+\pysr is a genetic algorithm-style symbolic regression algorithm
+\citep[see][for an early example]{kozaGeneticProgrammingMeans1994}.
+Loosely inspired by \textit{eureqa},
+\pysr searches the space of trees 
+containing a basis set of specified operators, variables, and arbitrary real constants
+to find an expression that best matches a given dataset.
+\pysr uses ``regularized evolution'' for its search strategy 
+\cite{realRegularizedEvolutionImage2019,realAutoMLZeroEvolvingMachine2020},
+which itself is a variant on classic tournament selection
+\cite{brindleGeneticAlgorithmsFunction1980,goldbergComparativeAnalysisSelection1991}.
+This algorithm simulates natural selection: each expression is represented by a tree, corresponding to its ``genetic code.''
+Expressions are subsampled, and the fittest expression in each group (using a metric which combines accuracy and simplicity) is used to breed new mutated expressions, as well as cross-bred with other top expressions.
+\pysr adds several additional techniques onto these classic algorithms, such as explicit constant optimization using BFGS \cite{fletcherPracticalMethodsOptimization1988}, and an adaptive complexity penalty, while allowing the use of custom operators and losses which is important for this work.
+
+\paragraph{Implementation for GWs}
+% Describe reanalysizing the real data using the extracted expression
+% After obtaining a number of 
+Symbolic regression is an automated type of empirical model discovery---something which has driven the development of many theories in the physical sciences.
+For example, Kepler's laws were an empirical model crafted by hand to fit patterns in the orbits of planets, and ultimately inspired Newton to discover a formula for gravitational force which could explain Kepler's laws \citep[see][for a review of this history]{hawkingShouldersGiantsGreat2004}.
+Astronomy has seen increasing use of symbolic regression to find new empirical models in an automated way \citep[e.g.,][]{grahamMachineassistedDiscoveryRelationships2013,cranmerDiscoveringSymbolicModels2020,wadekarModelingAssemblyBias2020,delgadoModelingGalaxyhaloConnection2021,cranmerHistogramPoolingOperators2021,shaoFindingUniversalRelations2021}.
+
+Here, we use symbolic regression for the first time to learn population density models in astrophysics.
+One of the main advantages of symbolic regression is to learn a model into an interpretable language, and, therefore, we define a set of operators which frequently occur in GW population models: addition, subtraction, division, power laws, Gaussians (unary operator: $f(x)=\exp(-x^2)$), and conditional statements (binary operator: $g(x, y)=\operatorname{if}(x \geq 0, y, 0)$).
+
+\section{Gravitational-Wave Data}
+\label{sec:GWdata}
+
+GW data comes in the form of time series, in which each time sample carries very little physical meaning.
+It is easier to understand the physical implication of a GW event through its inferred source properties \cite{Veitch:2014wba}.
+In the case of a BBH, the source properties include the masses of the two merging black holes.
+Understanding the population of GW events is essentially understanding the distribution of source properties, such as the mass distribution of black holes \cite{2019PASA...36...10T,Vitale:2020aaz}.
+
+In principle, one can apply symbolic regression to fit these distributions directly.
+Unfortunately, GW observations are extremely noisy.
+As a result, all inferred properties about the source such as the observed masses or distance to the source come with sizable uncertainty, often comparable to the prior used in inferring their source properties.
+This also means the posterior distribution of these inferred properties often has a non-trivial shape that the typical Gaussian approximation of "error bar" does not hold.
+On top of that, since there are only 79 announced events, the shot noise makes it very difficult to extract meaningful information if we first stack the events and fit the distribution with symbolic regression.
+
+In order to maximally extract information from the GW catalog, the community has employed Hierarchical Bayesian analysis (HBA) to analyze the distribution of GW events.
+HBA takes advantage of the fact that measurement uncertainties for each event are conditionally independent from the population model, so we can deconvolve the measurement uncertainty for each event (\cite{10.1214/10-AOAS439} describes this as extreme deconvolution) while fitting for the underlying distribution of GW events\footnote{For interested readers, \cite{Mandel:2018mve,Vitale:2020aaz,2019MNRAS.484.4008G} are excellent references with practical examples on how to apply the HBA framework to GW data.
+}.
+
+\newcommand\ppd{posterior predictive distribution\xspace} %PPD\xspace
+\newcommand{\defineppd}{} % (\ppd)
+
+One product coming out of this HBA pipeline is the posterior predictive distribution \defineppd for the properties of interest, i.e.
+a set of probability density functions that encapsulate the inferred source distribution and the uncertainty on the population level.
+It is much easier to find an interpretable form of the source distribution by fitting the \ppd with SR.
+In this study, we focus on understanding the distribution of binary black hole primary mass, i.e. the heavier mass in the binary.
+One of the most flexible models that is fitted to this set of data is a Gaussian mixture model \cite{Tiwari:2020vym}.
+To generate the data that we feed to the SR pipeline,
+we select a number of evaluation points in the primary mass axis,
+then we take the median of the \ppd given by the Gaussian mixture model as the value we are trying to predict.
+We also estimate the uncertainty of the value at each evaluation point using the $68\%$ confidence interval.
+
+\section{Result}
+\label{sec:result}
+
+We show the workflow of this work in figure \ref{fig:Gaussian_m1_fit}.
+On the leftmost panel, we show the posterior density in primary mass for every event in GWTC3.
+In the middle panel, we show a Gaussian mixture model is fitted to the data given by \cite{Tiwari:2020vym}.
+Given the \ppd shown in the middle panel, we use symbolic regression to extract effective descriptions of the data. 
+As we increase the complexity of the equation, it is natural to expect that the equation should fit the data better than a lower complexity equation since it has more free parameters.
+At the same time, an equation with higher complexity could either be overfitting or difficult to interpret.
+\pysr uses a score (see \citealt{cranmerDiscoveringSymbolicModels2020} for details) as a metric normalized to complexity to show how well an equation is fitting the data.
+At each complexity, \pysr outputs the equation with the lowest loss.
+To select the equations shown in figure \ref{fig:Gaussian_m1_fit}, we first make an accuracy cut based on the loss of the function, so the selected equation would be a good fit to the data.
+Then under a certain loss level, we select the equation that has the highest score.
+We choose three accuracy levels to select three different equations in our analysis.
+The equations selected under these criteria are tabulated in table \ref{tab:equations}.
+Note that we take the absolute value of the equations in order to make figure \ref{fig:Gaussian_m1_fit}.
+
+The lowest complexity (16) equation returned is a decaying exponential function with a Gaussian bump at the lower end of the mass distribution.
+This agrees with a recent study pointing out there is a very strong excess of events at $m_{1}\sim 10\ M_{\odot}$.
+In order to account for the bump around $30\ M_{\odot}$,
+the exponential part of this equation has a much higher normalization compared to the others.
+This agrees with the analysis done in GWTC3, where a single power-law seems to have trouble accounting for the excess of events around $m_{1}\sim 30\ M_{\odot}$\cite{LIGOScientific:2021psn}.
+
+Compared to the lowest complexity equation, the middle complexity (25) equation instead prefers a lower overall normalization of the exponential function,
+but account for the bump around $30\ M_{\odot}$ with a Gaussian component.
+This equation is similar to one of the simplest phenomenological models fitted to GWTC3, which is a power-law mixed with a Gaussian bump.
+The difference between this equation and the one presented in GWTC3 is this equation has an additional bump at $m_{1}\sim 10\ M_{\odot}$,
+which is also present in the lower complexity equation.
+This suggests the first bump at lower mass has more statistical significance than the second bump.
+Although the middle complexity equation is more complex, the loss is about one order of magnitude lower than the next best one with lower complexity (16).
+As a result, the score of this equation is higher than the lower complexity equation, making it a favorable equation.
+This means the second Gaussian component is needed in order to describe the distribution of GWs.
+
+At the highest complexity (48), one more Gaussian component is added to the equation in order to account for the tapering at the higher mass end. 
+Interestingly, the new component added in this equation does not try to fit the wiggles in between the first bump and the second bump.
+The significance of any features other than the two bumps and the continuum is still an open question.
+Our result hints the wiggles between the bumps seem to be less significant than the extra tapering of the continuum.
+In addition, none of the equations prefers to fit the extra hump before the $10\ M_{\odot}$ peak.
+This could mean the little hump could be an artifact in the Gaussian mixture model instead of a truly statistically significant feature.
+However, the loss of this equation is more or less similar to the middle complexity (25) equation, at the same time being almost twice as complex.
+As a result, the score is much lower, which means the tapering is not necessarily needed.
+
+\section{Discussion}
+\label{sec:discussion}
+
+% Brief summary of the paper
+In this paper, we use symbolic regression to extract interpretable equations from the \ppd of a flexible model fitted to the catalog of GWs.
+We show SR is able to discover equations that are similar to state-of-the-art phenomenological models.
+On top of finding forms that have already been discovered,
+our result also shed insights on an open question---whether there are additional features other than the two prominent peaks in the primary mass distribution \cite{Tiwari:2021yvr}.
+Our results suggest any additional features such as tapering of the continuum or wiggles between the two peaks are not statistically significant.
+
+Compared to traditional methods such as phenomenological modeling, which relies on an individual's insight,
+our pipeline is fully automatic.
+This allows the community to discover interpretable patterns in the GW catalog in a data-driven instead of model-driven manner.
+Hence, our method is more scalable to more complex and higher-dimensional datasets.
+
+In this work, we focus on analyzing only the primary mass distribution.
+Our method is trivially generalizable to higher-dimensional data, such as the joint space of both the primary mass and the secondary mass.
+So far there are relatively few phenomenological models which can capture correlations between observables,
+mainly due to the difficulty of writing down a reasonable model to account for these correlations.
+At the same time, the most advanced flexible model in the literature is an n-dimensional histogram, which is hard to interpret.
+Our method can be a natural way to explore and to explain these correlations.
+
+% \bibliographystyle{icml2022}
+\bibliography{bib.bib}
 

### 2022-06-27 src/tex/ms.tex
Somehow the running title is off

@@
-%%%%%%%% ICML 2022 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-
 \documentclass[nohyperref]{article}
+\usepackage{aas_macros}
 \usepackage{showyourwork}
@@
 % Therefore, a short form for the running title is supplied here:
-%\icmltitlerunning{Automated discovery of interpretable gravitational-wave population models}
+
+\icmltitlerunning{Short}
 
@@
 \twocolumn[
-%\icmltitle{Automatic, Flexible and Interpretable:\\gravitational-wave population analysis with symbolic regression}
-%\icmltitle{Automated discovery of symbolic gravitational-wave population models}
 \icmltitle{Automated discovery of interpretable gravitational-wave population models}
 
-\icmlsetsymbol{equal}{*}
-
 \begin{icmlauthorlist}
-\icmlauthor{Kaze W.K Wong}{equal,yyy}
-\icmlauthor{Miles Cranmer}{equal,yyy,comp}
+\icmlauthor{Kaze W.K Wong}{flatiron}
+\icmlauthor{Miles Cranmer}{princeton}
 \end{icmlauthorlist}
@@
 \icmlaffiliation{flatiron}{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
+\icmlaffiliation{princeton}{Department of Astrophysical Sciences, Princeton University, Princeton, New Jersey 08544, USA}
 
@@
 \section{Introduction}
-% \begin{figure*}[hbt!]
-% % \includegraphics[width=\textwidth]{figures/GWTC3_m1_combined.pdf}
+\begin{figure*}[hbt!]
+% \includegraphics[width=\textwidth]{figures/GWTC3_m1_combined.pdf}
 % \includegraphics[width=\textwidth]{figures/Page2.pdf}
-% \caption{An illustration of our proposed pipeline, which first
-% (\textit{far left})
-% requires observations of individual source properties (here, events shown from GWTC3);
-% then
-% (\textit{middle left}) 
-% infers the deconvolved population posterior density;
-% followed by (\textit{middle right})
-% using symbolic regression to search for an
-% an analytic model which approximates the density, using
-% operators common to existing population models; and finally
-% (\textit{far right}) distilling physical insights
-% from the recovered population model.
-% }
-% %\caption{\textit{left}: posterior density of the primary mass from all events reported in GWTC3. Each line is the normalized posterior density for one event.
-% %\textit{middle}: posterior predictive distribution of the Gaussian mixture model reported in \cite{LIGOScientific:2021psn}.
-% %\textit{right}: equations tabulated in table \ref{tab:equations}}
-% \label{fig:Gaussian_m1_fit}
-% \end{figure*}
+\caption{An illustration of our proposed pipeline, which first
+(\textit{far left})
+requires observations of individual source properties (here, events shown from GWTC3);
+then
+(\textit{middle left}) 
+infers the deconvolved population posterior density;
+followed by (\textit{middle right})
+using symbolic regression to search for an
+an analytic model which approximates the density, using
+operators common to existing population models; and finally
+(\textit{far right}) distilling physical insights
+from the recovered population model.
+}
+%\caption{\textit{left}: posterior density of the primary mass from all events reported in GWTC3. Each line is the normalized posterior density for one event.
+%\textit{middle}: posterior predictive distribution of the Gaussian mixture model reported in \cite{LIGOScientific:2021psn}.
+%\textit{right}: equations tabulated in table \ref{tab:equations}}
+\label{fig:Gaussian_m1_fit}
+\end{figure*}
 The rate of gravitational wave (GW) detections increases exponentially as the sensitivity of GW detectors improves \cite{LIGOScientific:2021djp} since the discovery in 2015 \cite{Abbott:2016blz}.
@@
 
-% \bibliographystyle{icml2022}
-\bibliography{bib.bib}
+\bibliographystyle{icml2022}
+\bibliography{bib}
 

### 2022-07-09 src/tex/ms.tex
Hack style file to get the running head working
add figure as static
Add improved methods for SR

@@
 \usepackage{subfigure}
+\usepackage{natbib}
 \usepackage{booktabs} % for professional tables
@@
 % Use the following line for the initial blind version submitted for review:
-\usepackage{icml2022}
+% \usepackage{icml2022}
 
 % If accepted, instead use the following line for the camera-ready submission:
-% \usepackage[accepted]{icml2022}
+\usepackage[accepted]{icml2022}
 
@@
 % Therefore, a short form for the running title is supplied here:
-
-\icmltitlerunning{Short}
 
@@
 % \includegraphics[width=\textwidth]{figures/GWTC3_m1_combined.pdf}
-% \includegraphics[width=\textwidth]{figures/Page2.pdf}
+\includegraphics[width=\textwidth]{figures/Page2.pdf}
 \caption{An illustration of our proposed pipeline, which first
@@
     \hline
-    Complexity &  Score & Loss & Equation\\
+    Complexity &  Score & Loss & Expression\\
     \hline
@@
     \end{tabular}
-    \caption{Equations return obtained through symbolic regression with \pysr.
+    \caption{Expressions obtained through symbolic regression with \pysr.
     In the search we perform, there are 30 equations with different complexities.
@@
 \pysr adds several additional techniques onto these classic algorithms, such as explicit constant optimization using BFGS \cite{fletcherPracticalMethodsOptimization1988}, and an adaptive complexity penalty, while allowing the use of custom operators and losses which is important for this work.
+
+Internal to \pysr, a multi-objective optimization problem is solved: the evolution seeks to find expressions which are as accurate as possible, while being as simple as possible.
+Formally, for our setting, \pysr solves the following optimization problem:
+\begin{align}
+    s_c &= 
+        \mathop{\arg\min}_{
+            \tiny
+            \begin{array}{c}s\in\mathcal{S}\\ \abs{s} = c \end{array}
+        }
+        \mathcal{L}(s),
+        \\
+    &\text{for each } c \in \{1, \ldots, 80\},\nonumber\\
+    \text{where }\mathcal{L}(s) &= \sum_i \ell(s(\mathbf{x}_i), y_i)
+\end{align}
+where $s$ is an expression, 
+$\mathcal{S}$ is a space of possible expressions considered,
+$\abs{s}$ is the complexity of an expression $s$,
+$\ell$ is the loss function, and $\{(\mathbf x_i, y_i)\}$ is the dataset
+for $\mathbf x_i$ each input and $y_i$ the label.
+The most accurate expression at each $c$, denoted as $s_c$, is computed for a given set range of $c$, which we choose as $c\in\{1, \ldots, 80\}$.
+Complexity is defined here
+as a sum of the complexity of each symbol in the expression,
+and $\ell(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$.
+
+Next, given the set of the best expressions at each complexity $\mathcal{S}^\ast = \{s_c\}_{c\in\{1,\ldots,80\}}$,
+we can define the \textit{score} of an expression as the slope of the accuracy-complexity curve, equal to:
+\begin{equation}
+    \text{score} = -\frac{\Delta (\log \mathcal{L}(s))}{\Delta(\abs{s})},
+\end{equation}
+which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \cite{eureqa,cranmerDiscoveringSymbolicModels2020}.
+Finally, we can construct the \textit{Pareto curve}, defined as:
+\begin{equation}
+    \hat{\mathcal{S}}^\ast = \{s_c \ |\ \mathcal{L}(s_c) < \mathcal{L}(s_q) \forall c > q \},
+\end{equation}
+such that each remaining expression is more accurate than all simpler expressions.
 

### 2022-07-09 src/tex/ms.tex
Rename figure

--- 2022-06-27 src/tex/ms.tex
@@
 However, flexible models come with a cost: a large number of parameters that lack physical motivation, making interpreting the implication of these models very difficult.
-In this work, we demonstrate the use of symbolic regression to distill such flexible models into interpretable analytic expressions.
+In this work, we demonstrate symbolic regression can complement flexible models by distilling the posterior predictive distribution of such flexible models into interpretable analytic expressions.
 We recover common GW population models such as a power-law-plus-Gaussian, and find a new empirical population model which combines accuracy and simplicity.
@@
 % \includegraphics[width=\textwidth]{figures/GWTC3_m1_combined.pdf}
-% \includegraphics[width=\textwidth]{figures/Page2.pdf}
+\includegraphics[width=\textwidth]{static/Gaussian_m1_fit.pdf}
 \caption{An illustration of our proposed pipeline, which first
@@
 
+Note that the result from symbolic regression does not completely replace the flexible model for a few reasons.
+First, symbolic regression does not
+
 In this work, we focus on analyzing only the primary mass distribution.

### 2022-07-10 src/tex/ms.tex
Add emphasis

@@
 % \includegraphics[width=\textwidth]{figures/GWTC3_m1_combined.pdf}
-\includegraphics[width=\textwidth]{static/Gaussian_m1_fit.pdf}
+\includegraphics[width=\textwidth]{../static/Gaussian_m1_fit.pdf}
 \caption{An illustration of our proposed pipeline, which first
@@
 our result also shed insights on an open question---whether there are additional features other than the two prominent peaks in the primary mass distribution \cite{Tiwari:2021yvr}.
-Our results suggest any additional features such as tapering of the continuum or wiggles between the two peaks are not statistically significant.
+Our results suggest interpreting any additional features such as tapering of the continuum or wiggles between the two peaks can be tricky as their significance are unclear.
 
+The main message of this work is that by combining flexible models with symbolic regression, we can discover interpretable equations that can describe the increasingly complex GW catalog.
 Compared to traditional methods such as phenomenological modeling, which relies on an individual's insight,
-our pipeline is fully automatic.
+our pipeline requires minimal human intervene.
+Our method is not a model selection method by itself, but a pipeline to discover models for downstream tasks such as model selection and parameter estimation.
+Once we obtain a set of equations, we can plug the expressions in a standard HBA pipeline to estimate the parameters of the model and select between models.
 This allows the community to discover interpretable patterns in the GW catalog in a data-driven instead of model-driven manner.
-Hence, our method is more scalable to more complex and higher-dimensional datasets.
-
-Note that the result from symbolic regression does not completely replace the flexible model for a few reasons.
-First, symbolic regression does not
+In this sense, our method is more scalable to more complex and higher-dimensional datasets.
 

### 2022-07-10 src/tex/ms.tex
Add meaning to the two bumps

@@
 The lowest complexity (16) equation returned is a decaying exponential function with a Gaussian bump at the lower end of the mass distribution.
-This agrees with a recent study pointing out there is a very strong excess of events at $m_{1}\sim 10\ M_{\odot}$.
+This agrees with a recent study pointing out there is a very strong excess of events at $m_{1}\sim 10\ M_{\odot}$,
+which corresponds to a smooth turn-on of the black hole mass spectrum \cite{Talbot:2018cva}.
 In order to account for the bump around $30\ M_{\odot}$,
@@
 Compared to the lowest complexity equation, the middle complexity (25) equation instead prefers a lower overall normalization of the exponential function,
-but account for the bump around $30\ M_{\odot}$ with a Gaussian component.
+but account for the bump around $30\ M_{\odot}$ with a Gaussian component,
+which is proposed to be driven by a pile-up from pulsational pair-instability supernovae \cite{Talbot:2018cva}.
 This equation is similar to one of the simplest phenomenological models fitted to GWTC3, which is a power-law mixed with a Gaussian bump.

### 2022-07-10 src/tex/ms.tex
Fix formatting with earlier icml2022 import

--- 2022-07-09 src/tex/ms.tex
@@
 \documentclass[nohyperref]{article}
+% If accepted, instead use the following line for the camera-ready submission:
+\usepackage[accepted]{icml2022}
+
 \usepackage{aas_macros}
@@
 
-% If accepted, instead use the following line for the camera-ready submission:
-\usepackage[accepted]{icml2022}
-
 % For theorems and such
@@
 \usepackage{amsthm}
+\usepackage{physics}
 
@@
 However, flexible models come with a cost: a large number of parameters that lack physical motivation, making interpreting the implication of these models very difficult.
-In this work, we demonstrate symbolic regression can complement flexible models by distilling the posterior predictive distribution of such flexible models into interpretable analytic expressions.
+In this work, we demonstrate the use of symbolic regression to distill such flexible models into interpretable analytic expressions.
 We recover common GW population models such as a power-law-plus-Gaussian, and find a new empirical population model which combines accuracy and simplicity.
@@
 % \includegraphics[width=\textwidth]{figures/GWTC3_m1_combined.pdf}
-\includegraphics[width=\textwidth]{static/Gaussian_m1_fit.pdf}
+\includegraphics[width=\textwidth]{figures/Page2.pdf}
 \caption{An illustration of our proposed pipeline, which first
@@
 
-Note that the result from symbolic regression does not completely replace the flexible model for a few reasons.
-First, symbolic regression does not
-
 In this work, we focus on analyzing only the primary mass distribution.

### 2022-07-10 src/tex/ms.tex
Reformat methods

@@
 Next, given the set of the best expressions at each complexity $\mathcal{S}^\ast = \{s_c\}_{c\in\{1,\ldots,80\}}$,
-we can define the \textit{score} of an expression as the slope of the accuracy-complexity curve, equal to:
-\begin{equation}
-    \text{score} = -\frac{\Delta (\log \mathcal{L}(s))}{\Delta(\abs{s})},
-\end{equation}
-which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \cite{eureqa,cranmerDiscoveringSymbolicModels2020}.
-Finally, we can construct the \textit{Pareto curve}, defined as:
+Next, we can construct the \textit{Pareto curve}, defined as:
 \begin{equation}
@@
 such that each remaining expression is more accurate than all simpler expressions.
+Using this, we can define the \textit{score} of an expression as the slope of the Pareto curve, equal to:
+\begin{equation}
+    \text{score} = \frac{\Delta (-\log \mathcal{L}(s))}{\Delta(\abs{s})},
+\end{equation}
+which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \cite{eureqa,cranmerDiscoveringSymbolicModels2020}.
 

### 2022-07-11 src/tex/ms.tex
Adjust abstract

@@
 \begin{abstract}
-We present an approach to automatically discover analytic population models for gravitational-wave (GW) events from data.
+We present an automatic approach to discover analytic population models for gravitational-wave (GW) events from data.
 As more gravitational-wave (GW) events are detected, flexible models such as Gaussian Mixture Models have become more important in fitting the distribution of GW properties due to their expressivity.
-However, flexible models come with a cost: a large number of parameters that lack physical motivation, making interpreting the implication of these models very difficult.
+However, flexible models come with many parameters that lack physical motivation, making interpreting the implication of these models challenging.
 In this work, we demonstrate the use of symbolic regression to distill such flexible models into interpretable analytic expressions.
 We recover common GW population models such as a power-law-plus-Gaussian, and find a new empirical population model which combines accuracy and simplicity.
-This demonstrates a promising pathway to automatically discover interpretable descriptions of features in the ever-growing GW catalog. 
+This demonstrates a promising pathway to discover interpretable population models, automatically, from the GW merger catalogs and other astrophysical phenomena.
 \end{abstract}

### 2022-07-11 src/tex/ms.tex
Adjust figure caption

@@
 followed by (\textit{middle right})
-using symbolic regression to search for an
-an analytic model which approximates the density, using
-operators common to existing population models; and finally
+using symbolic regression to search for an analytic model which approximates the density, using
+operators common to existing population models; and, finally
 (\textit{far right}) distilling physical insights

### 2022-07-11 src/tex/ms.tex
Clean introduction

@@
 \end{figure*}
-The rate of gravitational wave (GW) detections increases exponentially as the sensitivity of GW detectors improves \cite{LIGOScientific:2021djp} since the discovery in 2015 \cite{Abbott:2016blz}.
-To date, there are 79 officially announced events, the majority of them being binary black holes (BBH).
-With the increasing number of observed GW events, more and more features in the distribution of GW events await to be discovered.
+The rate of gravitational wave (GW) detections is increasing exponentially as the sensitivity of GW detectors improves \cite{LIGOScientific:2021djp} since the discovery in 2015 \cite{Abbott:2016blz}.
+Today there exist 79 official events, the majority of them being Binary Black Holes (BBHs).
+As more GW events are observed, more features in the distribution of GW events become resolved and available for discovery.
 Currently, one of the most common ways to understand the observed GW population is through phenomenological modeling.
-For example, to understand the distribution of primary mass in observed BBHs, one would write down a hand-crafted model of the primary mass distribution, such as a power-law distribution \cite{LIGOScientific:2021psn}.
-By comparing the model to the data, one can measure the parameters which characterize the model, such as the spectral index of a power-law distribution.
+For example, to understand the distribution of primary mass in observed BBHs, one would hand-craft a model of the primary mass distribution, such as a power-law distribution \cite{LIGOScientific:2021psn}.
+By comparing the model to the data, one can measure the parameters which characterize the model, such as the spectral index.
+\todo{Could make a statement about how power laws typically have some astrophysical meaning - hence a good operator, and why they are used in population models.}
 
-The central issue with the use of this paradigm is the lack of scalability to a more complex dataset, since manual input is required to write down a sensible phenomenological model.
+The central issue with this paradigm is the lack of scalability to a more complex dataset, since construction of sensible phenomenological models requires manual intervention.
+\todo{Weird transition here.}
 After the announcement of 11 events in the first Gravitational-Wave Transient Catalog (GWTC1) in 2019 \cite{LIGOScientific:2018mvr}, a power law was sufficient to explain the primary mass distribution.
-In GWTC2 (which is announced in 2020) \cite{LIGOScientific:2020kqk}, the community modified the simple power-law to either a broken power law with two spectral indices or with the superimposing of a Gaussian distribution on top of the power-law.
+In GWTC2 (which is announced in 2020) \cite{LIGOScientific:2020kqk}, the community modified the simple power law to either a broken power law with two spectral indices, or with the addition of a Gaussian distribution.
 Finally, the state-of-the-art model used in analyzing GWTC3 \cite{LIGOScientific:2021psn} further adds a spline on top of the power-law and Gaussian model to fit the residual \cite{Edelman:2021zkw}.
 Since the complexity of the catalog grows with the size of the catalog,
-the community is struggling to write down an interpretable model that is flexible enough to explain the increasing rich set of features in the catalog.
+the community is looking to develop an interpretable model that is flexible enough to explain the increasing rich set of features in the catalog.
 
-Alternatively, there have been studies using much more flexible models such as a Gaussian mixture model to analyze the catalog \cite{Tiwari:2020vym}.
+Besides such physically motivated models, there have been studies using much more flexible models such as a Gaussian Mixture Model (GMM) to analyze the catalog \cite{Tiwari:2020vym}.
 While such a flexible model is powerful in fitting complex datasets,
 it is difficult to interpret the physical meaning of the fitted parameters.
-There are two reasons why flexible models are hard to interpret:
-first, flexible models often have many parameters.
-In GWTC3, the specific Gaussian mixture model has over 100 parameters.
-With that number of parameters, it is not trivial to assess the relative importance of the parameters.
-Second, flexible models often employ bases that are not physically meaningful.
-For example, consider that one wishes to assess which parameter in the fitted Gaussian mixture model is the most important, one may apply a principal component analysis (PCA) to the set of parameters and find the most important component.
+There are two reasons why such flexible models are hard to interpret:
+first, these models often have many parameters;
+in GWTC3, the GMM has over 100 parameters.
+With so many degrees of freedom, it is non-trivial to analyze contributions from each mode and attempt to disentangle meaningful patterns.
+Second, flexible models often employ bases that are not physically motivated.
+For example, consider if one wished to assess which parameter in a fitted GMM is the most important: one may apply a Principal Component Analysis (PCA) to the set of parameters and find the most important component.
 However, the PCA basis for this dataset is not a basis that is directly related to any astrophysical process,
-so having the relative importance in that form does not provide insight into astrophysics.
+so having the relative importance in that form does not provide insight into astrophysics. \todo{This makes no sense. Wouldn’t you look at the amplitude?}
 
-Symbolic regression (SR) is a machine learning technique which searches to space of analytic expressions to fit a dataset---this algorithm complements the weakness of a flexible model in understanding the distribution of GW.
-Since SR does not take advantage of using gradient information when searching for expressions,
-and the space of equation is a combinatorial space,
-SR is slow in comparison to traditional machine learning techniques such as Gaussian Mixture Models (GMMs).
-Therefore, rather than fitting SR directly to the complex and expensive likelihoods found in GW event analysis, we first fit the GMM as a sort of approximated likelihood.
-Then, we apply SR to the output of this GMM, as the loss function is then simply mean-squared error compared to the GMM, rather equal to an evaluation of the raw likelihood.
+Symbolic regression (SR) is a machine learning technique which searches to space of analytic expressions to fit a dataset---this algorithm is both flexible and interpretable.
+SR can fit  many different patterns, while at the same time presenting the pattern in a simple concise way in the language of existing models—i.e., analytic expressions.
+SR complements the weakness of a flexible model in understanding the distribution of GW.
+However, since SR does not take advantage of using gradient information when searching for expressions,
+and  SR searches a combinatorial space over expressions,
+it is slow in comparison to traditional machine learning techniques such as Gaussian Mixture Models (GMMs).
+Therefore, rather than fitting SR directly to the complex and expensive likelihoods found in GW event analysis, we first fit the GMM as a fast-to-evaluate approximation of the true posterior.
+Then, we run SR to approximate the output of this GMM, which is feasible only because the loss function is then simply mean-squared error against the GMM, rather equal to an evaluation of the raw  and expensive likelihood.
 %While SR itself is not efficient enough to be used to analyze the data directly, using SR to interpret a flexible model that has already been fitted to the data can be very effective.
@@
 In this work, we apply SR to the mixture Gaussian model used to fit the primary mass distribution of GWTC3.
-In particular, we apply SR to fit the posterior predictive distribution given by the Gaussian mixture model.
+Specifically, we run SR to approximate the posterior predictive distribution produced by the Gaussian mixture model.
 This paper is structured as follows: we describe the GW data used in this work in \ref{sec:GWdata}.

### 2022-07-11 src/tex/ms.tex
Clean SR methods

@@
 
-Symbolic regression is a machine learning (ML) algorithm for regression which searches the space of analytic expressions that fit some dataset.
-Whereas most ML algorithms are optimized by fitting a fixed set number of real parameters, symbolic regression is typically done using a ``genetic algorithm'' which finds a tree or sequence of symbols that optimizes an objective.
-Since models for physical phenomena are often specified in a language of concise mathematical expressions, symbolic regression can produce very interpretable models compared to traditional ML, as the ``trained model'' generated by symbolic regression is in the same language as existing physical models.
+Symbolic regression is a machine learning (ML) algorithm for regression which searches  for analytic expressions which optimize some loss function.
+Whereas one can fit most ML algorithms to data by optimizing a fixed set number of real parameters, for symbolic regression, one typically uses using a ``genetic algorithm.'' This genetic algorithm arranges symbols in a tree, evaluates the resultant expression, and repeats this process until a particular goal is met.
+Since models for physical phenomena are often specified in a language of concise mathematical expressions, symbolic regression can produce very interpretable models compared to traditional ML, as the ``trained model'' generated by symbolic regression is in the same \textit{language} as existing physical models (i.e., analytic expressions).
 Therefore, symbolic regression can produce insights for science in a way that traditional ML cannot.

### 2022-07-11 src/tex/ms.tex
Clean table caption

@@
     In the search we perform, there are 30 equations with different complexities.
-    We select three most representative equations using a combination of score and loss threshold.}
+    We select three most representative equations by setting loss thresholds, and selecting the highest scoring expression below each threshold.
+This strategy rules out trivial expressions with low accuracy which had high scores only because they were adjacent to other trivial expressions.}
     \label{tab:equations}

### 2022-07-11 src/tex/ms.tex
Various edits throughout results

@@
 containing a basis set of specified operators, variables, and arbitrary real constants
-to find an expression that best matches a given dataset.
+to find expressions which fit relations in a given dataset.
 \pysr uses ``regularized evolution'' for its search strategy 
@@
 Expressions are subsampled, and the fittest expression in each group (using a metric which combines accuracy and simplicity) is used to breed new mutated expressions, as well as cross-bred with other top expressions.
-\pysr adds several additional techniques onto these classic algorithms, such as explicit constant optimization using BFGS \cite{fletcherPracticalMethodsOptimization1988}, and an adaptive complexity penalty, while allowing the use of custom operators and losses which is important for this work.
+\pysr adds several additional techniques onto these classic algorithms, such as explicit constant optimization using the \textit{BFGS} optimization algorithm \cite{fletcherPracticalMethodsOptimization1988}, and an adaptive complexity penalty, while allowing the use of custom operators and losses which is important for this work.
 
@@
 as a sum of the complexity of each symbol in the expression,
-and $\ell(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$.
+and $\ell(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$. \todo{What complexities were chosen?}
 
@@
 
+\todo{Note how much computational power was used, what other PySR settings were used. Could link the script in the paper repo.}
+
 \paragraph{Implementation for GWs}
@@
 One of the main advantages of symbolic regression is to learn a model into an interpretable language, and, therefore, we define a set of operators which frequently occur in GW population models: addition, subtraction, division, power laws, Gaussians (unary operator: $f(x)=\exp(-x^2)$), and conditional statements (binary operator: $g(x, y)=\operatorname{if}(x \geq 0, y, 0)$).
+\todo{Specify how combinations of operators are ruled out}
 
@@
 In the case of a BBH, the source properties include the masses of the two merging black holes.
-Understanding the population of GW events is essentially understanding the distribution of source properties, such as the mass distribution of black holes \cite{2019PASA...36...10T,Vitale:2020aaz}.
+Understanding the population of GW events supports out understanding of the distribution of source properties, such as the mass distribution of black holes \cite{2019PASA...36...10T,Vitale:2020aaz}.
 
@@
 As a result, all inferred properties about the source such as the observed masses or distance to the source come with sizable uncertainty, often comparable to the prior used in inferring their source properties.
-This also means the posterior distribution of these inferred properties often has a non-trivial shape that the typical Gaussian approximation of "error bar" does not hold.
-On top of that, since there are only 79 announced events, the shot noise makes it very difficult to extract meaningful information if we first stack the events and fit the distribution with symbolic regression.
+This also means the posterior distribution of these inferred properties often has a non-trivial shape that the typical Gaussian approximation of ``error bar’’ does not hold.
+Moreover, since there are only 79 announced events, the shot noise makes it challenging to extract meaningful information if we first stack the events and fit the distribution with symbolic regression.
 
-In order to maximally extract information from the GW catalog, the community has employed Hierarchical Bayesian analysis (HBA) to analyze the distribution of GW events.
+In order to maximally extract information from the GW catalog, the community has employed Hierarchical Bayesian Analysis (HBA) to analyze the distribution of GW events.
 HBA takes advantage of the fact that measurement uncertainties for each event are conditionally independent from the population model, so we can deconvolve the measurement uncertainty for each event (\cite{10.1214/10-AOAS439} describes this as extreme deconvolution) while fitting for the underlying distribution of GW events\footnote{For interested readers, \cite{Mandel:2018mve,Vitale:2020aaz,2019MNRAS.484.4008G} are excellent references with practical examples on how to apply the HBA framework to GW data.
@@
 
-One product coming out of this HBA pipeline is the posterior predictive distribution \defineppd for the properties of interest, i.e.
-a set of probability density functions that encapsulate the inferred source distribution and the uncertainty on the population level.
+One product coming out of this HBA pipeline is the posterior predictive distribution \defineppd for the properties of interest, i.e., a set of probability density functions that encapsulate the inferred source distribution and the uncertainty on the population level.
 It is much easier to find an interpretable form of the source distribution by fitting the \ppd with SR.
-In this study, we focus on understanding the distribution of binary black hole primary mass, i.e. the heavier mass in the binary.
+In this study, we focus on understanding the distribution of binary black hole primary mass, i.e., the heavier mass in the binary.
 One of the most flexible models that is fitted to this set of data is a Gaussian mixture model \cite{Tiwari:2020vym}.
 To generate the data that we feed to the SR pipeline,
-we select a number of evaluation points in the primary mass axis,
+we select many evaluation points in the primary mass axis,
 then we take the median of the \ppd given by the Gaussian mixture model as the value we are trying to predict.
@@
 On the leftmost panel, we show the posterior density in primary mass for every event in GWTC3.
-In the middle panel, we show a Gaussian mixture model is fitted to the data given by \cite{Tiwari:2020vym}.
+In the middle panel \todo{Shouldn’t it be left-middle?}, we show a GMM fitted to the data given by \cite{Tiwari:2020vym}.
 Given the \ppd shown in the middle panel, we use symbolic regression to extract effective descriptions of the data. 
-As we increase the complexity of the equation, it is natural to expect that the equation should fit the data better than a lower complexity equation since it has more free parameters.
+As we increase the complexity of the equation, it is natural to expect that the equation should fit the data better than a lower complexity equation since it has additional free parameters.
 At the same time, an equation with higher complexity could either be overfitting or difficult to interpret.
@@
 Then under a certain loss level, we select the equation that has the highest score.
-We choose three accuracy levels to select three different equations in our analysis.
+We choose three accuracy levels to select three different equations in our analysis \todo{Specify how the thresholds were chosen}.
 The equations selected under these criteria are tabulated in table \ref{tab:equations}.
-Note that we take the absolute value of the equations in order to make figure \ref{fig:Gaussian_m1_fit}.
+Note that we take the absolute value of the equations in order to make figure \ref{fig:Gaussian_m1_fit}.\todo{Is this still true?}
 
@@
 the exponential part of this equation has a much higher normalization compared to the others.
-This agrees with the analysis done in GWTC3, where a single power-law seems to have trouble accounting for the excess of events around $m_{1}\sim 30\ M_{\odot}$\cite{LIGOScientific:2021psn}.
+This agrees with the analysis done in GWTC3, where a single power law seems to have trouble accounting for the excess of events around $m_{1}\sim 30\ M_{\odot}$\cite{LIGOScientific:2021psn}.
 
-Compared to the lowest complexity equation, the middle complexity (25) equation instead prefers a lower overall normalization of the exponential function,
+Compared to the lowest complexity equation, the middle complexity (25) equation instead prefers a lower normalization of the exponential function,
 but account for the bump around $30\ M_{\odot}$ with a Gaussian component.
-This equation is similar to one of the simplest phenomenological models fitted to GWTC3, which is a power-law mixed with a Gaussian bump.
+This equation is similar to one of the simplest phenomenological models fitted to GWTC3, which is a power law mixed with a Gaussian bump.
 The difference between this equation and the one presented in GWTC3 is this equation has an additional bump at $m_{1}\sim 10\ M_{\odot}$,
@@
 This could mean the little hump could be an artifact in the Gaussian mixture model instead of a truly statistically significant feature.
-However, the loss of this equation is more or less similar to the middle complexity (25) equation, at the same time being almost twice as complex.
+However, the loss of this equation is similar to the middle complexity (25) equation, at the same time being almost twice as complex.
 As a result, the score is much lower, which means the tapering is not necessarily needed.
@@
 In this paper, we use symbolic regression to extract interpretable equations from the \ppd of a flexible model fitted to the catalog of GWs.
-We show SR is able to discover equations that are similar to state-of-the-art phenomenological models.
+We show SR can discover equations which are similar to state-of-the-art phenomenological models.
 On top of finding forms that have already been discovered,
-our result also shed insights on an open question---whether there are additional features other than the two prominent peaks in the primary mass distribution \cite{Tiwari:2021yvr}.
-Our results suggest any additional features such as tapering of the continuum or wiggles between the two peaks are not statistically significant.
+our result also shed insights on an open question---whether there exist additional features other than the two prominent peaks in the primary mass distribution \cite{Tiwari:2021yvr}.
+Our results suggest any additional features such as tapering of the continuum or wiggles between the two peaks are not statistically significant \todo{I would lighten the language of this part - SR isn’t exactly statistically rigorous}.
 
@@
 So far there are relatively few phenomenological models which can capture correlations between observables,
-mainly due to the difficulty of writing down a reasonable model to account for these correlations.
+mainly due to the difficulty of manually constructing reasonable models to account for these correlations.
 At the same time, the most advanced flexible model in the literature is an n-dimensional histogram, which is hard to interpret.
-Our method can be a natural way to explore and to explain these correlations.
+Our method can be a natural way to explore and to explain these correlations. \todo{Say something about how this can be used for other astrophysical phenomena!}
 

### 2022-07-11 src/tex/ms.tex

@@
 However, flexible models come with many parameters that lack physical motivation, making interpreting the implication of these models challenging.
-In this work, we demonstrate the use of symbolic regression to distill such flexible models into interpretable analytic expressions.
+In this work, we demonstrate symbolic regression can complement flexible models by distilling the posterior predictive distribution of such flexible models into interpretable analytic expressions.
 We recover common GW population models such as a power-law-plus-Gaussian, and find a new empirical population model which combines accuracy and simplicity.
@@
 % \includegraphics[width=\textwidth]{figures/GWTC3_m1_combined.pdf}
-\includegraphics[width=\textwidth]{figures/Page2.pdf}
+\includegraphics[width=\textwidth]{../static/Gaussian_m1_fit.pdf}
 \caption{An illustration of our proposed pipeline, which first
@@
 Compared to the lowest complexity equation, the middle complexity (25) equation instead prefers a lower normalization of the exponential function,
-but account for the bump around $30\ M_{\odot}$ with a Gaussian component.
+but account for the bump around $30\ M_{\odot}$ with a Gaussian component,
+which is proposed to be driven by a pile-up from pulsational pair-instability supernovae \cite{Talbot:2018cva}.
 This equation is similar to one of the simplest phenomenological models fitted to GWTC3, which is a power law mixed with a Gaussian bump.
@@
 our result also shed insights on an open question---whether there exist additional features other than the two prominent peaks in the primary mass distribution \cite{Tiwari:2021yvr}.
-Our results suggest any additional features such as tapering of the continuum or wiggles between the two peaks are not statistically significant \todo{I would lighten the language of this part - SR isn’t exactly statistically rigorous}.
+Our results suggest interpreting any additional features such as tapering of the continuum or wiggles between the two peaks can be tricky as their significance are unclear. \todo{Clean up language}
 
+The main message of this work is that by combining flexible models with symbolic regression, we can discover interpretable equations that can describe the increasingly complex GW catalog.
 Compared to traditional methods such as phenomenological modeling, which relies on an individual's insight,
-our pipeline is fully automatic.
+our pipeline requires minimal human intervene.
+Our method is not a model selection method by itself, but a pipeline to discover models for downstream tasks such as model selection and parameter estimation.
+Once we obtain a set of equations, we can plug the expressions in a standard HBA pipeline to estimate the parameters of the model and select between models.
 This allows the community to discover interpretable patterns in the GW catalog in a data-driven instead of model-driven manner.
-Hence, our method is more scalable to more complex and higher-dimensional datasets.
+In this sense, our method is more scalable to more complex and higher-dimensional datasets.
 

### 2022-07-11 src/tex/ms.tex
Change up to Symbolic regression section

@@
 We recover common GW population models such as a power-law-plus-Gaussian, and find a new empirical population model which combines accuracy and simplicity.
-This demonstrates a promising pathway to discover interpretable population models, automatically, from the GW merger catalogs and other astrophysical phenomena.
+This demonstrates a promising pathway to automatically discover interpretable population models in the ever-growing GW catalog.
 \end{abstract}
@@
 The rate of gravitational wave (GW) detections is increasing exponentially as the sensitivity of GW detectors improves \cite{LIGOScientific:2021djp} since the discovery in 2015 \cite{Abbott:2016blz}.
-Today there exist 79 official events, the majority of them being Binary Black Holes (BBHs).
-As more GW events are observed, more features in the distribution of GW events become resolved and available for discovery.
+The most update catalog of GW contains 79 events, the majority of them being Binary Black Holes (BBHs).
+As more GW events are observed, more features in the distribution of GW events await to be discovered.
 Currently, one of the most common ways to understand the observed GW population is through phenomenological modeling.
@@
 By comparing the model to the data, one can measure the parameters which characterize the model, such as the spectral index.
-\todo{Could make a statement about how power laws typically have some astrophysical meaning - hence a good operator, and why they are used in population models.}
 
-The central issue with this paradigm is the lack of scalability to a more complex dataset, since construction of sensible phenomenological models requires manual intervention.
-\todo{Weird transition here.}
-After the announcement of 11 events in the first Gravitational-Wave Transient Catalog (GWTC1) in 2019 \cite{LIGOScientific:2018mvr}, a power law was sufficient to explain the primary mass distribution.
+The main issue with this paradigm is the lack of scalability to a more complex dataset, since construction of sensible phenomenological models requires manual intervention.
+To illustrate the shortcoming of this paradigm: 
+when 11 events were announced in the first Gravitational-Wave Transient Catalog (GWTC1) in 2019 \cite{LIGOScientific:2018mvr}, a power law was sufficient to explain the primary mass distribution.
 In GWTC2 (which is announced in 2020) \cite{LIGOScientific:2020kqk}, the community modified the simple power law to either a broken power law with two spectral indices, or with the addition of a Gaussian distribution.
@@
 Since the complexity of the catalog grows with the size of the catalog,
-the community is looking to develop an interpretable model that is flexible enough to explain the increasing rich set of features in the catalog.
+it becomes increasingly difficult to develop an interpretable model that is flexible enough to explain the increasing rich set of features in the catalog .
 
-Besides such physically motivated models, there have been studies using much more flexible models such as a Gaussian Mixture Model (GMM) to analyze the catalog \cite{Tiwari:2020vym}.
+Alternatively, there have been studies using much more flexible models such as a Gaussian Mixture Model (GMM) to analyze the catalog \cite{Tiwari:2020vym}.
 While such a flexible model is powerful in fitting complex datasets,
@@
 However, the PCA basis for this dataset is not a basis that is directly related to any astrophysical process,
-so having the relative importance in that form does not provide insight into astrophysics. \todo{This makes no sense. Wouldn’t you look at the amplitude?}
+so having the relative importance in that form does not provide insight into astrophysics directly.
+\todo{This makes no sense.
+Wouldn’t you look at the amplitude?} \kw{PCA is a transformed coordinate that describe the data, I can't look at the amplitude of a PCA component and know exactly what it means astrophysically.
+You won't report the PCA value in your abstract since it has no physical meaning.}
 
-Symbolic regression (SR) is a machine learning technique which searches to space of analytic expressions to fit a dataset---this algorithm is both flexible and interpretable.
-SR can fit  many different patterns, while at the same time presenting the pattern in a simple concise way in the language of existing models—i.e., analytic expressions.
-SR complements the weakness of a flexible model in understanding the distribution of GW.
+Symbolic regression (SR) is a machine learning technique which searches to space of analytic expressions to fit a dataset, which complements the weakness of a flexible model in understanding the distribution of GW.
 However, since SR does not take advantage of using gradient information when searching for expressions,
@@
 In this work, we apply SR to the mixture Gaussian model used to fit the primary mass distribution of GWTC3.
-Specifically, we run SR to approximate the posterior predictive distribution produced by the Gaussian mixture model.
+Specifically, we approximate the posterior predictive distribution produced of the Gaussian mixture model with SR.
 This paper is structured as follows: we describe the GW data used in this work in \ref{sec:GWdata}.
@@
 
+While most ML algorithms optimizie a fixed set number of real parameters,
 Symbolic regression is a machine learning (ML) algorithm for regression which searches for analytic expressions which optimize some loss function.
-Whereas one can fit most ML algorithms to data by optimizing a fixed set number of real parameters, for symbolic regression, one typically uses using a ``genetic algorithm.'' This genetic algorithm arranges symbols in a tree, evaluates the resultant expression, and repeats this process until a particular goal is met.
-Since models for physical phenomena are often specified in a language of concise mathematical expressions, symbolic regression can produce very interpretable models compared to traditional ML, as the ``trained model'' generated by symbolic regression is in the same \textit{language} as existing physical models (i.e., analytic expressions).
+The search typically employs a ``genetic algorithm'', which arranges symbols in a tree, evaluates the resultant expression, and repeats this process until a particular goal is met.
+Since models for physical phenomena are often specified in a language of concise mathematical expressions, symbolic regression can produce interpretable models compared to traditional ML, as the ``trained model'' generated by symbolic regression is in the same \textit{language} as existing physical models (i.e., analytic expressions).
 Therefore, symbolic regression can produce insights for science in a way that traditional ML cannot.
-
-% Give history
-%Symbolic regression has a disconnected history,
-% with some early work going back to \mc{langley citations}.
-%The first application of genetic algorithms
-%to symbolic regression was done by \mc{koza}, before this class of algorithms was popularized in the sciences by \mc{lipson}.
-% \mc{Cite examples in astronomy.}
-% \mc{Describe how PySR works.}
 

### 2022-07-11 src/tex/ms.tex
Address flexible model shortcoming and change abstract

@@
 We recover common GW population models such as a power-law-plus-Gaussian, and find a new empirical population model which combines accuracy and simplicity.
-This demonstrates a promising pathway to automatically discover interpretable population models in the ever-growing GW catalog.
+This demonstrates a promising pathway to automatically discover interpretable population models in the ever-growing GW catalog, which can potentially be applied to other astrophysical phenomena.
 \end{abstract}
@@
 Second, flexible models often employ bases that are not physically motivated.
-For example, consider if one wished to assess which parameter in a fitted GMM is the most important: one may apply a Principal Component Analysis (PCA) to the set of parameters and find the most important component.
-However, the PCA basis for this dataset is not a basis that is directly related to any astrophysical process,
-so having the relative importance in that form does not provide insight into astrophysics directly.
-\todo{This makes no sense.
-Wouldn’t you look at the amplitude?} \kw{PCA is a transformed coordinate that describe the data, I can't look at the amplitude of a PCA component and know exactly what it means astrophysically.
-You won't report the PCA value in your abstract since it has no physical meaning.}
+For example, the GMM models has 11 independent Gaussian components in the mass-spin space.
+While one or two Gaussians in mass or spin space alone can be physically motivated,
+the rest of the components are included in the model jsut to fit for unexpected features.
+Therefore, interpreting the importance of each component could be complicated by the components that are not physically motiviated.
 

### 2022-07-11 src/tex/ms.tex
Last part remain

@@
 as a sum of the complexity of each symbol in the expression,
-and $\ell(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$. \todo{What complexities were chosen?}
+and $\ell(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$.
+% \todo{What complexities were chosen?}
 
@@
 
-\todo{Note how much computational power was used, what other PySR settings were used. Could link the script in the paper repo.}
+% \todo{Note how much computational power was used, what other PySR settings were used. Could link the script in the paper repo.}
 
@@
 One of the main advantages of symbolic regression is to learn a model into an interpretable language, and, therefore, we define a set of operators which frequently occur in GW population models: addition, subtraction, division, power laws, Gaussians (unary operator: $f(x)=\exp(-x^2)$), and conditional statements (binary operator: $g(x, y)=\operatorname{if}(x \geq 0, y, 0)$).
-\todo{Specify how combinations of operators are ruled out}
+% \todo{Specify how combinations of operators are ruled out}
 
@@
 In the case of a BBH, the source properties include the masses of the two merging black holes.
-Understanding the population of GW events supports out understanding of the distribution of source properties, such as the mass distribution of black holes \cite{2019PASA...36...10T,Vitale:2020aaz}.
+On top of extracting physics from individual events,
+we can also learn about astrophysical processes through understanding of the distribution of source properties, such as the mass distribution of black holes \cite{2019PASA...36...10T,Vitale:2020aaz}.
 
@@
 On the leftmost panel, we show the posterior density in primary mass for every event in GWTC3.
-In the middle panel \todo{Shouldn’t it be left-middle?}, we show a GMM fitted to the data given by \cite{Tiwari:2020vym}.
+In the left-middle panel, we show a GMM fitted to the data given by \cite{Tiwari:2020vym}.
 Given the \ppd shown in the middle panel, we use symbolic regression to extract effective descriptions of the data. 
@@
 Then under a certain loss level, we select the equation that has the highest score.
-We choose three accuracy levels to select three different equations in our analysis \todo{Specify how the thresholds were chosen}.
+We choose three accuracy levels to select three different equations in our analysis.
+% \todo{Specify how the thresholds were chosen}
 The equations selected under these criteria are tabulated in table \ref{tab:equations}.
-Note that we take the absolute value of the equations in order to make figure \ref{fig:Gaussian_m1_fit}.\todo{Is this still true?}
+Note that we take the absolute value of the equations in order to make figure \ref{fig:Gaussian_m1_fit}.
 
@@
 our result also shed insights on an open question---whether there exist additional features other than the two prominent peaks in the primary mass distribution \cite{Tiwari:2021yvr}.
-Our results suggest interpreting any additional features such as tapering of the continuum or wiggles between the two peaks can be tricky as their significance are unclear. \todo{Clean up language}
+The additional suggest interpreting any additional features such as tapering of the continuum or wiggles between the two peaks can be tricky as their significance are unclear.
 
@@
 At the same time, the most advanced flexible model in the literature is an n-dimensional histogram, which is hard to interpret.
-Our method can be a natural way to explore and to explain these correlations. \todo{Say something about how this can be used for other astrophysical phenomena!}
+Our method can be a natural way to explore and to explain these correlations.
 

### 2022-07-11 src/tex/ms.tex

@@
 our result also shed insights on an open question---whether there exist additional features other than the two prominent peaks in the primary mass distribution \cite{Tiwari:2021yvr}.
-The additional suggest interpreting any additional features such as tapering of the continuum or wiggles between the two peaks can be tricky as their significance are unclear.
+Since the highest complexity equation does not significantly improve the fitting accuracy, one should take caution when interpreting the tapering of the continuum or wiggles between the two peaks as their significance are unclear.
 

### 2022-07-11 src/tex/ms.tex
Update ms.tex

@@
 
+Although we focus on GW in this work, the method can be easily extended to other population analysis.
+For example, given the recently release Gaia DR3 data \cite{2016A&A...595A...1G}, , we can apply our method to understanding the population of stars without much augmentation to the pipeline.
+As size of data sets grows larger in astronomy, this work provides a simple and scalable pathway to uncover trends and relations in astrophysical catalogs.
+
 \bibliographystyle{icml2022}

### 2022-07-11 src/tex/ms.tex
Update ms.tex

@@
 Although we focus on GW in this work, the method can be easily extended to other population analysis.
-For example, given the recently release Gaia DR3 data \cite{2016A&A...595A...1G}, , we can apply our method to understanding the population of stars without much augmentation to the pipeline.
+For example, given the recently release Gaia DR3 data \cite{Collaboration2016TheGM}, , we can apply our method to understanding the population of stars without much augmentation to the pipeline.
 As size of data sets grows larger in astronomy, this work provides a simple and scalable pathway to uncover trends and relations in astrophysical catalogs.

### 2022-07-11 src/tex/ms.tex
Update ms.tex

--- 2022-07-11 src/tex/ms.tex
@@
 Although we focus on GW in this work, the method can be easily extended to other population analysis.
-For example, given the recently release Gaia DR3 data \cite{2016A&A...595A...1G}, , we can apply our method to understanding the population of stars without much augmentation to the pipeline.
+For example, given the recently release Gaia DR3 data \cite{Collaboration2016TheGM}, we can apply our method to understanding the population of stars without much augmentation to the pipeline.
 As size of data sets grows larger in astronomy, this work provides a simple and scalable pathway to uncover trends and relations in astrophysical catalogs.

### 2022-07-11 src/tex/ms.tex
Fix citation

@@
 \end{equation}
-which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \cite{eureqa,cranmerDiscoveringSymbolicModels2020}.
+which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \cite{schmidtDistillingFreeFormNatural2009,cranmerDiscoveringSymbolicModels2020}.
 

### 2022-07-11 src/tex/ms.tex
Clean up text

@@
 
-We use the algorithm \pysr\footnote{\url{https://github.com/MilesCranmer/PySR}} \cite{cranmerPySRFastParallelized2020} for performing symbolic regression.
-\pysr is a genetic algorithm-style symbolic regression algorithm
+We use the algorithm and software \pysr\footnote{\url{https://github.com/MilesCranmer/PySR}} \cite{cranmerPySRFastParallelized2020} for performing symbolic regression.
+\pysr uses a genetic algorithm-style symbolic regression algorithm
 \citep[see][for an early example]{kozaGeneticProgrammingMeans1994}.
@@
 $\abs{s}$ is the complexity of an expression $s$,
-$\ell$ is the loss function, and $\{(\mathbf x_i, y_i)\}$ is the dataset
-for $\mathbf x_i$ each input and $y_i$ the label.
+$\ell$ is the loss function, and $\{(x_i, y_i)\}$ is the dataset
+for $x_i$ each input and $y_i$ the label.
 The most accurate expression at each $c$, denoted as $s_c$, is computed for a given set range of $c$, which we choose as $c\in\{1, \ldots, 80\}$.

### 2022-07-11 src/tex/ms.tex
Clean up C definition

@@
         \mathcal{L}(s),
-        \\
-    &\text{for each } c \in \{1, \ldots, 80\},\nonumber\\
-    \text{where }\mathcal{L}(s) &= \sum_i \ell(s(\mathbf{x}_i), y_i)
+    \quad\text{for each } c\in \mathcal{C},\nonumber\\
+    &\text{where }\mathcal{L}(s) = \sum_i \ell(s(\mathbf{x}_i), y_i)
 \end{align}
@@
 for $x_i$ each input and $y_i$ the label.
-The most accurate expression at each $c$, denoted as $s_c$, is computed for a given set range of $c$, which we choose as $c\in\{1, \ldots, 80\}$.
+The most accurate expression at each $c$, denoted as $s_c$, is computed for a given set range of $c$, which we choose as $c\in\mathcal{C}\equiv \{1, \ldots, 50\}$.
 Complexity is defined here

### 2022-07-11 src/tex/ms.tex
Give definitions of complexity

@@
 Complexity is defined here
-as a sum of the complexity of each symbol in the expression,
-and $\ell(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$.
-% \todo{What complexities were chosen?}
+as the number of symbols in the expression (including repeats),
+and $\ell$ is defined as $(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$.
 

### 2022-07-11 src/tex/ms.tex
Split title cleaner
Fix some typos

@@
 \twocolumn[
-\icmltitle{Automated discovery of interpretable gravitational-wave population models}
+\icmltitle{
+Automated discovery \\
+of interpretable gravitational-wave population models}
 
@@
 
-While most ML algorithms optimizie a fixed set number of real parameters,
-Symbolic regression is a machine learning (ML) algorithm for regression which searches for analytic expressions which optimize some loss function.
+While most ML algorithms optimize a fixed set number of real parameters,
+Symbolic regression is a machine learning (ML) algorithm which searches for analytic expressions which optimize some loss function.
 The search typically employs a ``genetic algorithm'', which arranges symbols in a tree, evaluates the resultant expression, and repeats this process until a particular goal is met.

### 2022-07-11 src/tex/ms.tex
Space out rows of table
Start discussing nested constraints

@@
     \hline
-    16 & 0.126   & 0.212  & $3.72 \rm{Cond}{\left(M - 9.27,0.9^{M} \right)} + 3.72 \rm{Gauss}{\left(0.52 M - 4.85 \right)}$\\
+    16 & 0.126   & 0.212  & $3.72 \rm{Cond}{\left(M - 9.27,0.9^{M} \right)} + 3.72 \rm{Gauss}{\left(0.52 M - 4.85 \right)}$\\[0.4em]
+    % Add more space between rows of table^
     25 & 0.501   & 0.0352 & $3.89 \left(\rm{Gauss}{\left(0.19 M - 6.54 \right)} + 0.45\right) \rm{Cond}{\left(M - 9.12,0.91^{M} \right)}$\\
-    & & & $ + 3.89 \rm{Gauss}{\left(0.5 M - 4.8 \right)}$\\
+    & & & $ + 3.89 \rm{Gauss}{\left(0.5 M - 4.8 \right)}$\\[0.4em]
     48 & 0.00771 & 0.0108 & $\left(\rm{Cond}{\left(M - 9.42,0.62 \cdot 0.9^{M} \right)} + 1.44 \rm{Gauss}{\left(0.51 M - 4.88 \right)}\right)$\\
@@
 Here, we use symbolic regression for the first time to learn population density models in astrophysics.
-One of the main advantages of symbolic regression is to learn a model into an interpretable language, and, therefore, we define a set of operators which frequently occur in GW population models: addition, subtraction, division, power laws, Gaussians (unary operator: $f(x)=\exp(-x^2)$), and conditional statements (binary operator: $g(x, y)=\operatorname{if}(x \geq 0, y, 0)$).
+One of the main advantages of symbolic regression is to learn a model into an interpretable language, and, therefore, we define a set of operators which frequently occur in GW population models: addition, subtraction, division, power laws, square, Gaussians (unary operator: $f(x)=\exp(-x^2)$), and conditional statements (binary operator: $g(x, y)=\operatorname{if}(x \geq 0, y, 0)$).
+To simplify our assumptions, all operators are assigned the same complexity of $1$.
+We also introduce additional constraints on how operators may be nested inside each other to rule out uninterpretable expressions; these are defined in \cref{sec:constraints}.
+
 % \todo{Specify how combinations of operators are ruled out}
@@
 
+\appendix
+\section{Operator constraints}
+\label{sec:constraints}
+
+We implement several constraints on the space of expressions, to rule out uninterpretable combinations of operators.
+We impose the following nested constraints:
+\begin{enumerate}
+    \item No operator with arity of 1 (i.e., taking a single argument) may be nested with another arity-1 operator, unless the operator is square.
+    \begin{enumerate}
+        \item However, a square can only be nested inside itself once.
+    \end{enumerate}
+    \item A division operator may only be nested inside itself once.
+    \item A power operator may not contain any arity-1 operator, except a square.
+\end{enumerate}
+
 \end{document}

### 2022-07-11 src/tex/ms.tex
Describe rest of constraints

@@
 
+\clearpage
 \appendix
@@
 
-We implement several constraints on the space of expressions, to rule out uninterpretable combinations of operators.
+With full flexibility over how symbolic operators can be arranged up to a maximum number, surprisingly many combinations exist which are simple and accurate, yet completely uninterpretable.
+Typically, a genetic algorithm might find that a particular deeply nested combination of operators might perform very well, with the overall complexity being low.
+However, we wish to find a model that is interpretable, and so we provide some additional constraints here.
+
+First, we implement limits on which operators can be nested, and how deeply.
 We impose the following nested constraints:
 \begin{enumerate}
-    \item No operator with arity of 1 (i.e., taking a single argument) may be nested with another arity-1 operator, unless the operator is square.
+    \item No operator with arity of 1 (i.e., an operator taking a single argument) may be nested with another arity-1 operator, unless the operator is square.
     \begin{enumerate}
-        \item However, a square can only be nested inside itself once.
+        \item However, we permit a square to be nested inside itself once, as such a combination may be used to make a quadratic operation with few operations.
     \end{enumerate}
-    \item A division operator may only be nested inside itself once.
-    \item A power operator may not contain any arity-1 operator, except a square.
+    \item The division operator may only be nested inside itself once.
+    \item The power operator may not contain any arity-1 operator.
 \end{enumerate}
+We also impose the following constraints on maximum complexity within each operator:
+\begin{enumerate}
+    \item The power operator may only have a maximum complexity of 9 in its base, and 3 in its exponent.
+    \item The division operator may have any complexity in the numerator, but a maximum complexity of 7 in the denominator.
+    \item The Gauss operator may only have a maximum complexity of 11 in its argument.
+    \item The conditional operator may only have a maximum complexity of 5 in its condition, and 11 in its resultant argument.
+\end{enumerate}
+These prevent \pysr from finding expressions which, despite being low-complexity and high-accuracy, are still uninterpretable to a human scientist due to unusual combinations of operators compared to typical expressions found throughout physics and astrophysics.
 

### 2022-07-11 src/tex/ms.tex
Add units to table
Correct selection strategy

@@
     \hline
-    Complexity &  Score & Loss & Expression\\
+    Complexity &  Score & Loss & Expression for $\text{Rate}[\text{yr}^{-1}](M=M/M_\odot)$\\
     \hline
@@
     In the search we perform, there are 30 equations with different complexities.
-    We select three most representative equations by setting loss thresholds, and selecting the highest scoring expression below each threshold.
-This strategy rules out trivial expressions with low accuracy which had high scores only because they were adjacent to other trivial expressions.}
+    We select three representative equations from the Pareto front by setting three successive complexity ranges, and selecting the highest scoring expression in each range.}
     \label{tab:equations}
@@
         \mathcal{L}(s),
-    \quad\text{for each } c\in \mathcal{C},\nonumber\\
-    &\text{where }\mathcal{L}(s) = \sum_i \ell(s(\mathbf{x}_i), y_i)
+    \quad\text{for each } c\in \mathcal{C},\nonumber
 \end{align}
+where the loss function $\mathcal{L}$ is defined as:
+\begin{equation}
+\mathcal{L}(s) = \sum_i \ell(s(\mathbf{x}_i), y_i),
+\end{equation}
 where $s$ is an expression, 

### 2022-07-11 src/tex/ms.tex
Fix naming of Pareto front

@@
 Next, given the set of the best expressions at each complexity $\mathcal{S}^\ast = \{s_c\}_{c\in\{1,\ldots,80\}}$,
-Next, we can construct the \textit{Pareto curve}, defined as:
+Next, we can construct the \textit{Pareto front}, defined as:
 \begin{equation}
-    \hat{\mathcal{S}}^\ast = \{s_c \ |\ \mathcal{L}(s_c) < \mathcal{L}(s_q) \forall c > q \},
+    \hat{\mathcal{S}}^\ast = \{s_c \ |\ \mathcal{L}(s_c) < \mathcal{L}(s_q) \forall q < c \},
 \end{equation}
 such that each remaining expression is more accurate than all simpler expressions.
-Using this, we can define the \textit{score} of an expression as the slope of the Pareto curve, equal to:
+Using this, we can define the \textit{score} of an expression as the slope of the Pareto front, equal to:
 \begin{equation}
@@
 \end{equation}
-which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \cite{schmidtDistillingFreeFormNatural2009,cranmerDiscoveringSymbolicModels2020}.
+which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \citep[see, e.g.,][]{schmidtDistillingFreeFormNatural2009,cranmerDiscoveringSymbolicModels2020}.
 

### 2022-07-11 src/tex/ms.tex
Specify complexity ranges

@@
 
-% \todo{Specify how combinations of operators are ruled out}
-
 \section{Gravitational-Wave Data}
@@
 At each complexity, \pysr outputs the equation with the lowest loss.
-To select the equations shown in figure \ref{fig:Gaussian_m1_fit}, we first make an accuracy cut based on the loss of the function, so the selected equation would be a good fit to the data.
-Then under a certain loss level, we select the equation that has the highest score.
-We choose three accuracy levels to select three different equations in our analysis.
-% \todo{Specify how the thresholds were chosen}
+To select the equations shown in figure \ref{fig:Gaussian_m1_fit}, we first make an accuracy cut based on the loss of the function such that all expressions have a loss lower than $0.3$, so the selected equation would still be a good fit to the data.
+We then consider the expressions with the highest score among three ranges: $c\in [1, 20]$,
+$c\in[20, 35]$, and $c\in[35, 50]$.
+These ranges are set to simply give us a variety of complexity expressions to study.
 The equations selected under these criteria are tabulated in table \ref{tab:equations}.
-Note that we take the absolute value of the equations in order to make figure \ref{fig:Gaussian_m1_fit}.
 

### 2022-07-11 src/tex/ms.tex
Additional text cleanup
Inline score definition to get more space

@@
         \mathcal{L}(s),
-    \quad\text{for each } c\in \mathcal{C},\nonumber
+    \quad\text{for each } c\in \mathcal{C},
 \end{align}
-where the loss function $\mathcal{L}$ is defined as:
-\begin{equation}
-\mathcal{L}(s) = \sum_i \ell(s(\mathbf{x}_i), y_i),
-\end{equation}
+for the loss function $\mathcal{L}(s) = \sum_i \ell(s(\mathbf{x}_i), y_i),$
 where $s$ is an expression, 
@@
 such that each remaining expression is more accurate than all simpler expressions.
-Using this, we can define the \textit{score} of an expression as the slope of the Pareto front, equal to:
-\begin{equation}
-    \text{score} = \frac{\Delta (-\log \mathcal{L}(s))}{\Delta(\abs{s})},
-\end{equation}
+Using this, we can define the \textit{score} of an expression as the slope of the Pareto front, equal to: $\text{score} = -\Delta(\log \mathcal{L}(s_c))/\Delta c,$
 which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \citep[see, e.g.,][]{schmidtDistillingFreeFormNatural2009,cranmerDiscoveringSymbolicModels2020}.
@@
 In order to maximally extract information from the GW catalog, the community has employed Hierarchical Bayesian Analysis (HBA) to analyze the distribution of GW events.
-HBA takes advantage of the fact that measurement uncertainties for each event are conditionally independent from the population model, so we can deconvolve the measurement uncertainty for each event (\cite{10.1214/10-AOAS439} describes this as extreme deconvolution) while fitting for the underlying distribution of GW events\footnote{For interested readers, \cite{Mandel:2018mve,Vitale:2020aaz,2019MNRAS.484.4008G} are excellent references with practical examples on how to apply the HBA framework to GW data.
+HBA takes advantage of the fact that measurement uncertainties for each event are conditionally independent from the population model, so we can deconvolve the measurement uncertainty for each event (\citealt{10.1214/10-AOAS439} describes this as extreme deconvolution) while fitting for the underlying distribution of GW events\footnote{For interested readers, \cite{Mandel:2018mve,Vitale:2020aaz,2019MNRAS.484.4008G} are excellent references with practical examples on how to apply the HBA framework to GW data.
 }.

### 2022-07-17 src/scripts/figure1_subparts.py
add data and plotting script

@@ -1,0 +1,138 @@
+import matplotlib.pyplot as plt
+from pysr import *
+import numpy as np
+import sympy
+import pandas as pd
+from sympy import preorder_traversal, Float
+
+plt.rcParams['font.family'] = 'serif'
+plt.rcParams['font.serif'] = 'Computer Modern Raman'
+plt.rcParams['text.usetex'] = True
+
+def get_data():
+    data = np.load("GWTC3_m1_spectrum.npz")
+    X = data["axis"]
+
+    X = np.concatenate(
+        [
+            X,
+            np.array([1e-9, 1e9]),
+        ]
+    )
+
+    # Samples: [5, 50, 95]
+    y_med = np.log(data["pm1_med"])
+    y_low = np.log(data["pm1_low"])
+    y_high = np.log(data["pm1_high"])
+
+    y_med = np.concatenate([y_med, np.array([-8, -8])])
+    y_low = np.concatenate([y_low, np.array([-12, -12])])
+    y_high = np.concatenate([y_high, np.array([-6, -6])])
+
+    stdev = (y_high - y_low) / 4
+    snr = 1 / stdev**2
+
+    X = pd.DataFrame(
+        {
+            "logM": np.log(X),
+            "M": X,
+            "logP": y_med,
+            "snr": snr,
+            "stdev": stdev,
+            "logP_med": y_med,
+            "logP_low": y_low,
+            "logP_high": y_high,
+        }
+    )
+    return X
+
+def round_expression(ex1):
+    ex2 = ex1
+    for a in preorder_traversal(ex1):
+        if isinstance(a, Float):
+            ex2 = ex2.subs(a, round(a, 2))
+    return ex2
+
+Gauss = lambda x: sympy.exp(-(x**2))
+Cond = lambda x, y: sympy.Piecewise((0, x < 0), (y, True))
+model = PySRRegressor(
+        binary_operators=["+", "-", "*", "pow", "cond(x,y)= x >= 0 ? y : 0", "custompow(x,y) = x > 0 ? x^y : -100"],
+    unary_operators=["square", "gauss(x)=exp(-x^2)"],
+    equation_file="multinode_search_gwaves_with_nested_constraints4.txt",
+    extra_sympy_mappings={
+        "gauss": lambda x: sympy.exp(-(x**2)),
[…]

### 2022-07-17 src/tex/ms.tex

@@
 We recover common GW population models such as a power-law-plus-Gaussian, and find a new empirical population model which combines accuracy and simplicity.
-This demonstrates a promising pathway to automatically discover interpretable population models in the ever-growing GW catalog, which can potentially be applied to other astrophysical phenomena.
+This demonstrates a strategy to automatically discover interpretable population models in the ever-growing GW catalog, which can potentially be applied to other astrophysical phenomena.
 \end{abstract}
@@
 
-Alternatively, there have been studies using much more flexible models such as a Gaussian Mixture Model (GMM) to analyze the catalog \cite{Tiwari:2020vym}.
+Alternatively, there exist studies using much more flexible models such as  Gaussian Mixture Models (GMMs) to analyze the catalog \cite{Tiwari:2020vym}.
 While such a flexible model is powerful in fitting complex datasets,
 it is difficult to interpret the physical meaning of the fitted parameters.
-There are two reasons why such flexible models are hard to interpret:
+There stand two core reasons why such flexible models are hard to interpret:
 first, these models often have many parameters;
@@
 While one or two Gaussians in mass or spin space alone can be physically motivated,
-the rest of the components are included in the model jsut to fit for unexpected features.
-Therefore, interpreting the importance of each component could be complicated by the components that are not physically motiviated.
+the additional components are included in the model fit for unexpected features.
+Therefore, interpreting the importance of each component could be complicated by the components that are not physically motivated.
 
@@
 While most ML algorithms optimize a fixed set number of real parameters,
-Symbolic regression is a machine learning (ML) algorithm which searches for analytic expressions which optimize some loss function.
-The search typically employs a ``genetic algorithm'', which arranges symbols in a tree, evaluates the resultant expression, and repeats this process until a particular goal is met.
+symbolic regression is a machine learning (ML) algorithm which searches for analytic expressions which optimize some loss function.
+The search typically employs a ``genetic algorithm,’’ which arranges symbols in a tree, evaluates the resultant expression, and repeats this process until a particular target is met.
 Since models for physical phenomena are often specified in a language of concise mathematical expressions, symbolic regression can produce interpretable models compared to traditional ML, as the ``trained model'' generated by symbolic regression is in the same \textit{language} as existing physical models (i.e., analytic expressions).
@@
 as the number of symbols in the expression (including repeats),
-and $\ell$ is defined as $(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$.
+and $\ell$ as $\ell(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$.
 
 Next, given the set of the best expressions at each complexity $\mathcal{S}^\ast = \{s_c\}_{c\in\{1,\ldots,80\}}$,
-Next, we can construct the \textit{Pareto front}, defined as:
+Next, we can construct the \textit{Pareto front}:
 \begin{equation}
@@
 $c\in[20, 35]$, and $c\in[35, 50]$.
-These ranges are set to simply give us a variety of complexity expressions to study.
+These ranges are set to give us a variety of complexity expressions to study.
 The equations selected under these criteria are tabulated in table \ref{tab:equations}.
@@
 As a result, the score of this equation is higher than the lower complexity equation, making it a favorable equation.
-This means the second Gaussian component is needed in order to describe the distribution of GWs.
 
-At the highest complexity (48), one more Gaussian component is added to the equation in order to account for the tapering at the higher mass end. 
+At the highest complexity (48), one more Gaussian component is added to the equation account for the tapering at the higher mass end. 
 Interestingly, the new component added in this equation does not try to fit the wiggles in between the first bump and the second bump.
@@
 The main message of this work is that by combining flexible models with symbolic regression, we can discover interpretable equations that can describe the increasingly complex GW catalog.
-Compared to traditional methods such as phenomenological modeling, which relies on an individual's insight,
-our pipeline requires minimal human intervene.
-Our method is not a model selection method by itself, but a pipeline to discover models for downstream tasks such as model selection and parameter estimation.
-Once we obtain a set of equations, we can plug the expressions in a standard HBA pipeline to estimate the parameters of the model and select between models.
-This allows the community to discover interpretable patterns in the GW catalog in a data-driven instead of model-driven manner.
+Compared to traditional methods such as phenomenological modeling, which relies on an individual scientist's insight,
+our pipeline requires minimal human intervention.
+ This technique is not a model selection method by itself, but a pipeline to discover models for downstream tasks such as model selection and parameter estimation.
+Once we obtain a set of equations, we can insert the expressions to a standard HBA pipeline to estimate the parameters of the model and select between models.
+This allows the community to discover interpretable patterns in the GW catalog in a data-driven instead of hand-crafted model-driven manner.
 In this sense, our method is more scalable to more complex and higher-dimensional datasets.
@@
 At the same time, the most advanced flexible model in the literature is an n-dimensional histogram, which is hard to interpret.
-Our method can be a natural way to explore and to explain these correlations.
+Our method provides a natural way to explore and to explain these correlations.
 
-Although we focus on GW in this work, the method can be easily extended to other population analysis.
-For example, given the recently release Gaia DR3 data \cite{Collaboration2016TheGM}, we can apply our method to understanding the population of stars without much augmentation to the pipeline.
-As size of data sets grows larger in astronomy, this work provides a simple and scalable pathway to uncover trends and relations in astrophysical catalogs.
+Although we focus on GW in this work, this method can be easily extended to other population analysis problems.
+For example, given the recently released Gaia DR3 dataset \cite{Collaboration2016TheGM}, it might be possible to apply this method to understanding various subpopulations of  stars in the Milky Way.
+As size of datasets grows larger in astronomy, this work provides a simple and scalable pathway to uncover trends and relations in astrophysical catalogs.
 
@@
 \appendix
-\section{Operator constraints}
+\section{Operator Constraints}
 \label{sec:constraints}

### 2022-07-19 src/scripts/figure1_subparts.py
Move site to base

@@ -1,138 +1,0 @@
-import matplotlib.pyplot as plt
-from pysr import *
-import numpy as np
-import sympy
-import pandas as pd
-from sympy import preorder_traversal, Float
-
-plt.rcParams['font.family'] = 'serif'
-plt.rcParams['font.serif'] = 'Computer Modern Raman'
-plt.rcParams['text.usetex'] = True
-
-def get_data():
-    data = np.load("GWTC3_m1_spectrum.npz")
-    X = data["axis"]
-
-    X = np.concatenate(
-        [
-            X,
-            np.array([1e-9, 1e9]),
-        ]
-    )
-
-    # Samples: [5, 50, 95]
-    y_med = np.log(data["pm1_med"])
-    y_low = np.log(data["pm1_low"])
-    y_high = np.log(data["pm1_high"])
-
-    y_med = np.concatenate([y_med, np.array([-8, -8])])
-    y_low = np.concatenate([y_low, np.array([-12, -12])])
-    y_high = np.concatenate([y_high, np.array([-6, -6])])
-
-    stdev = (y_high - y_low) / 4
-    snr = 1 / stdev**2
-
-    X = pd.DataFrame(
-        {
-            "logM": np.log(X),
-            "M": X,
-            "logP": y_med,
-            "snr": snr,
-            "stdev": stdev,
-            "logP_med": y_med,
-            "logP_low": y_low,
-            "logP_high": y_high,
-        }
-    )
-    return X
-
-def round_expression(ex1):
-    ex2 = ex1
-    for a in preorder_traversal(ex1):
-        if isinstance(a, Float):
-            ex2 = ex2.subs(a, round(a, 2))
-    return ex2
-
-Gauss = lambda x: sympy.exp(-(x**2))
-Cond = lambda x, y: sympy.Piecewise((0, x < 0), (y, True))
-model = PySRRegressor(
-        binary_operators=["+", "-", "*", "pow", "cond(x,y)= x >= 0 ? y : 0", "custompow(x,y) = x > 0 ? x^y : -100"],
-    unary_operators=["square", "gauss(x)=exp(-x^2)"],
-    equation_file="multinode_search_gwaves_with_nested_constraints4.txt",
-    extra_sympy_mappings={
-        "gauss": lambda x: sympy.exp(-(x**2)),
[…]

### 2022-07-19 src/scripts/paths.py

@@ -1,26 +1,0 @@
-"""
-Exposes common paths useful for manipulating datasets and generating figures.
-
-"""
-from pathlib import Path
-
-# Absolute path to the top level of the repository
-root = Path(__file__).resolve().parents[2].absolute()
-
-# Absolute path to the `src` folder
-src = root / "src"
-
-# Absolute path to the `src/data` folder (contains datasets)
-data = src / "data"
-
-# Absolute path to the `src/static` folder (contains static images)
-static = src / "static"
-
-# Absolute path to the `src/scripts` folder (contains figure/pipeline scripts)
-scripts = src / "scripts"
-
-# Absolute path to the `src/tex` folder (contains the manuscript)
-tex = src / "tex"
-
-# Absolute path to the `src/tex/figures` folder (contains figure output)
-figures = tex / "figures"

### 2022-07-20 src/scripts/figure1_subparts.py
Update plotting scripts

@@ -5,9 +5,29 @@ import sympy
 import pandas as pd
 from sympy import preorder_traversal, Float
 
+
 plt.rcParams['font.family'] = 'serif'
 plt.rcParams['font.serif'] = 'Computer Modern Raman'
 plt.rcParams['text.usetex'] = True
+
+GWTC = np.load('./src/data/GWTC3_posterior.npz')
+m1 = GWTC['m1']
+from scipy.stats import gaussian_kde
+m1_axis = np.linspace(0,100,1000)
+fig, ax = plt.subplots(1,1,figsize=(10,9))
+for i in range(79):
+    kde = gaussian_kde(m1[i,::10])
+    ax.plot(m1_axis,kde(m1_axis),c='grey',alpha=0.5)
+ax.set_yscale('log')
+ax.set_ylim(1e-3,3)
+ax.set_xlim(0,100)
+ax.set_xlabel(r'$m_{1,i}\ [M_{\odot}]$',fontsize=20)
+ax.set_ylabel(r'$p(m_{1,i})$',fontsize=20)
+ax.tick_params(axis='x',labelsize=16)
+ax.tick_params(axis='y',labelsize=16)
+ax.set_title('Data',fontsize=24)
+fig.savefig('./figure/GWTC3_m1_data.pdf',bbox_inches='tight')
+
 
 def get_data():
     data = np.load("GWTC3_m1_spectrum.npz")
@@ -118,13 +138,24 @@ label3 = label3.replace('&','\&')
 
 
 
-fig, ax = plt.subplots(1,1,figsize=(20,9))
-ax.plot(X['M'].values[:-2],np.exp(X['logP'])[:-2],label='Gaussian Mixture Data',lw=5)
+fig, ax = plt.subplots(1,1,figsize=(10,9))
+ax.plot(X['M'].values[:-2],np.exp(X['logP'])[:-2],lw=5)
 ax.fill_between(X['M'].values[:-2],np.exp(X['logP_low'][:-2]),np.exp(X['logP_high'][:-2]),alpha=0.14,color='C0')
-ax.plot(X['M'].values[:-2],np.abs(y1[:-2]),label='PySR,'+label1,lw=3)
-ax.plot(X['M'].values[:-2],np.abs(y2[:-2]),label='PySR,'+label2,lw=3)
[…]
